---
title: "hw6"
author: "sze pui"
date: "11/28/2021"
output: github_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
```

#Problem 1
##tidy data
lets import and clean the library first...
For  tidying data we have mainly two tasks: 
1.convert numeric to factor where appropriate
2.check for missing data

```{r}
#task 1 convert numeric to factor where appropriate
birthweight_df =read_csv("data/birthweight.csv") %>% 
  janitor::clean_names()%>% 
  #convert babysex,frace,malform,mrace into factor and recode them 
  mutate( babysex= as.factor(babysex),
          frace= as.factor(frace),
          malform = as.factor(malform),
          mrace = as.factor(mrace))%>% 
  #babysex: baby’s sex (male = 1, female = 2)
  mutate(babysex=  recode(babysex,"1" = "male","2"= "female"),
         frace= recode(frace,"1" = "white", "2" = "black", "3" = "asian", "4" = "puerto rican", "8" = "other", "9" = "unknown"),
        malform=  recode(malform,"0" = "absent", "1" = "present" ),
        mrace = recode(mrace,"1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other")) %>% 
                           rename(dad_race = frace, mom_race = mrace)
  
#task 2  check for missing data
sum(is.na(birthweight_df))
```
There is no missing data in the birthweight dataframe, therefore we dont habe to deal with the missing data using such as na.omit()/drop_na()

##Regression model
Propose a regression model for birthweight.

When I run the linear regression with all variables, I only keep the variables showing a very small p-values. A small p value of a variable indicates that it is significant to the baby's birthweight.For example, menarche ,mother’s age at menarche, is eliminated since it has a large p-value with 2.261406e-01	
Therefore, based on the above selection rule, I selected the following variables with very small varibales: baby’s head circumference , baby’s length, mother’s weight,babysex, family monthly income, gestational age,mom's age,  mother’s pre-pregnancy BMI, average number of cigarettes smoked per day during pregnancy
```{r}
#Model fitting
fit = lm(bwt ~ blength +bhead + delwt+ babysex +fincome +gaweeks +momage +ppbmi +smoken , data = birthweight_df) 
 
#tidying output 
fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)%>% 
  knitr::kable(digits = 3)


```
Let's do the diagnostics!
```{r}
birthweight_df %>% 
  modelr::add_residuals(fit) %>% 
   modelr::add_predictions(fit) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point()+
   labs(x ="predicted baby’s birth weight (grams)", 
        y= "residuals", 
        title = "scatterplots of prediction against residuals ")
```



#Comparision to two other models 
